Robust Control of a Fluxonium Qubit

### 2020-12-04 paper/5static.tex
updates

\todo{Ask Brian to review this paragraph and next two}
The sampling method simultaneously optimizes over copies of a state,
each of which evolves with a distinct deviant parameter value.
We append four sets of two sample states $\psi_{\pm}$
to the augmented state vector \eqref{eq:astatecontrols}.
We initialize each pair of samples in the same state.
The four initial states are chosen so that their outerproducts
span the operators on the Hilbert space
$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
The discrete dynamics \eqref{eq:dyn_con} is modified so that the
samples evolve under the Hamiltonian given in \eqref{eq:hamiltonian}
with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$ for a fixed $\sigma_{f_{q}}$.
The final state
for each sample is the image of its initial state
under the desired gate. We penalize deviations
from the final state in the objective \eqref{eq:costfun}.

For the unscented sampling method, we append $2(2n + d)$ samples
to the augmented state vector \eqref{eq:astatecontrols}
for each initial state in the operator basis. Here $2n = 4$ is twice the
dimension of the Hilbert space, resulting from the isomorphism \eqref{eq:isomorphism},
and $d = 1$ is the number of deviant parameters. The samples
represent a Gaussian distribution over the $2n$
elements of the initial state, modeling
the uncertainty in the state as a result of the uncertainty in
the deviant parameter. Each sample evolves under the Hamiltonian
given in \eqref{eq:hamiltonian} with $f_{q} \gets f_{q} + \delta f_{q}$.
We determine the qubit frequency detuning $\delta f_{q}$
at each knot point using the statistics of the samples.
We modify the discrete dynamics \eqref{eq:dyn_con} to
evolve the samples under the deviant Hamiltonian
and then apply the unscented transformation to the ensemble
of samples,
accurately preserving the first and second moments
of the distribution.
As in the sampling method,
we penalize deviations from the final state according to \eqref{eq:costfun}.
A detailed procedure for the unscented transformation is given
in Appendix \ref{appendix:unscented}.

The derivative method penalizes the sensitivty of the state
to the deviant parameter, which is encoded in the ($l$\textsuperscript{th}-order)
derivative $\partial_{f_{q}}^{l} \psi$. In the $m$\textsuperscript{th}-order
derivative method, we append all state derivatives of order $1, \dots, m$
to the augmented state vector \eqref{eq:astatecontrols}
for each initial state in the operator basis.
We penalize the norms of the state derivatives
in \eqref{eq:costfun} by setting the corresponding final states to zero.
We could obtain the state derivatives in the discrete dynamics \eqref{eq:dyn_con}
with backwards mode differentiation.
Naive automatic differentiation 
would differentiate all dynamics at knot points
$1, \dots, k - 1$ to obtain the state derivative at knot point $k$, requiring
$O(N^{2})$ matrix multiplications. Instead, we 
employ forwards mode differentiation on the TDSE to obtain coupled dynamics
which require $O(N)$ matrix multiplications to integrate.
For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
  H \ket{\partial_{f_{q}} \psi} +
  (\partial_{f_{q}} H) \ket{\psi}
  \label{eq:d1dyn}
\end{align}
Exponential integrators that account for the non-linear
term in \eqref{eq:d1dyn} are used to efficiently integrate the coupled dynamics,
see Appendix \ref{appendix:derivative}. For timing
results and complexity analyses for the three methods,
consult Appendix \ref{appendix:time}.

>>>>

\todo{Ask Brian to review this paragraph and next two}
The sampling method simultaneously optimizes over copies of a state,
each of which evolves with a distinct deviant parameter value.
We append four sets of two sample states $\psi^{\pm}$
to the augmented state vector \eqref{eq:astatecontrols}.
We initialize each pair of samples in the same state.
The four initial states are chosen so that their outerproducts
span the operators on the Hilbert space
$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
The discrete dynamics \eqref{eq:dyn_con} is modified so that the
samples evolve under the Hamiltonian given in \eqref{eq:hamiltonian}
with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$ for a fixed $\sigma_{f_{q}}$.
The final state
for each sample is the image of its initial state
under the desired gate. The infidelity between
the samples and their final states are penalized
by adding a term $\sum_{k=1}^{N} q_{k} (1 - \braket{\psi^{\pm}_{N}}{\psi_{f}})$
to the objective \eqref{eq:costfun}.

For the unscented sampling method, we append $2(2n + d)$ samples
to the augmented state vector \eqref{eq:astatecontrols}
for each initial state in the operator basis. Here $2n = 4$ is twice the
dimension of the Hilbert space, resulting from the isomorphism \eqref{eq:isomorphism},
and $d = 1$ is the number of deviant parameters. The samples
represent a Gaussian distribution over the $2n$
elements of the initial state, modeling
the uncertainty in the state as a result of the uncertainty in
the deviant parameter. Each sample evolves under the Hamiltonian
given in \eqref{eq:hamiltonian} with $f_{q} \gets f_{q} + \delta f_{q}$.
We determine the qubit frequency detuning $\delta f_{q}$
at each knot point using the statistics of the samples.
We modify the discrete dynamics \eqref{eq:dyn_con} to
evolve the samples under the deviant Hamiltonian
and then apply the unscented transformation to the ensemble
of samples,
accurately preserving the first and second moments
of the distribution.
As for the sampling method,
we penalize the infidelity between the samples and their final states
by adding an infidelity contribution to \eqref{eq:costfun}.
A detailed procedure for the unscented transformation is given
in Appendix \ref{appendix:unscented}.

The derivative method penalizes the sensitivty of the state
to the deviant parameter, which is encoded in the ($l$\textsuperscript{th}-order)
derivative $\partial_{f_{q}}^{l} \psi$. In the $m$\textsuperscript{th}-order
derivative method, we append all state derivatives of order $1, \dots, m$
to the augmented state vector \eqref{eq:astatecontrols}
for each initial state in the operator basis.
We penalize the norms of the state derivatives
in \eqref{eq:costfun} by setting the corresponding final states to zero.
We could obtain the state derivatives in the discrete dynamics \eqref{eq:dyn_con}
with backwards mode differentiation.
Naive automatic differentiation 
would differentiate all dynamics at knot points
$1, \dots, k - 1$ to obtain the state derivative at knot point $k$, requiring
$O(N^{2})$ matrix multiplications. Instead, we 
employ forwards mode differentiation on the TDSE to obtain coupled dynamics
which require $O(N)$ matrix multiplications to integrate.
For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
  H \ket{\partial_{f_{q}} \psi} +
  (\partial_{f_{q}} H) \ket{\psi}
  \label{eq:d1dyn}
\end{align}
Exponential integrators that account for the non-linear
term in \eqref{eq:d1dyn} are used to efficiently integrate the coupled dynamics,
see Appendix \ref{appendix:derivative}. For runtimes
and complexity analyses for the three methods,
consult Appendix \ref{appendix:time}.

### 2020-12-04 paper/a4time.tex
updates

\begin{table}[ht!]
  \begin{tabular} {c | c | c | c | c | c | c}
    & \multicolumn{6}{c}{Wall Time ($\pm 0.001$ s)}\\
    \hline
    $t_{N}$ (ns) & Base & Long. & S & SU & D1 & D2\\
    \hline
    $18$ & 0.155 & 1.688 & 1.773 & & 16.713 & 48.398\\ % 0.155049
    $36$ & 7.014 & - & 48.213 & & 67.838 & 81.030\\
    $72$ & 15.906 & - & 281.372 & & 266.997 & 332.182\\
  \end{tabular}
  \caption{
    \todo{finish timing table}
    Runtimes for $Z/2$ optimizations.
    Programs were executed on a single thread of an
    AMD Ryzen Threadripper 3970X 32-Core Processor.
  }
  \label{tab:time}
\end{table}

\todo{Ask Dave if we should keep complexity analysis in main text.}
Now we present the problem size complexities for the robustness methods.
For the sampling method, the size of the augmented state vector
is $O(dn^{3})$. $d$ is the number of deviant parameters.
$n$ is the dimension of the Hilbert space. There are $n^{2}$ initial states
in the operator basis, $2d$ sample states per initial state,
and each state has $2n$ elements.
For the unscented sampling method, the size of the augmented state vector
is $O(dn^{3} + n^{4})$.
There are $n^{2}$ initial states in the operator basis,
$2(2n + d)$ sample states per initial state,
and each state has $2n$ elements.
For the $m$\textsuperscript{th}-order derivative method, the size of the augmented state vector
is $O(dmn^{3})$. There are $n^{2}$
initial states in the operator basis, $dm$ state derivatives per initial state,
and each state has $2n$ elements.

>>>>

\begin{table}[ht!]
  \begin{tabular} {c | c | c | c | c | c | c}
    & \multicolumn{6}{c}{Wall Time ($\pm 0.001$ s)}\\
    \hline
    $t_{N}$ (ns) & Base & Long. & S & SU & D1 & D2\\
    \hline
    $18$ & 0.155 & 1.688 & 1.773 & 210.573 & 16.713 & 48.398\\ % 0.155049
    $36$ & 7.014 & - & 48.213 & 4566.236 & 67.838 & 81.030\\
    $72$ & 15.906 & - & 281.372 & 16575.308 & 266.997 & 332.182\\
  \end{tabular}
  \caption{
    Runtimes for $Z/2$ optimizations.
    Programs were executed on a single thread of an
    AMD Ryzen Threadripper 3970X 32-Core Processor.
  }
  \label{tab:time}
\end{table}

\todo{Ask Dave if we should keep complexity analysis in main text.}
Now we present the problem size complexities for the robustness methods.
For the sampling method, the size of the augmented state vector
is $O(dn^{3})$. $d$ is the number of deviant parameters.
$n$ is the dimension of the Hilbert space. There are $n^{2}$ initial states
in the operator basis, $2d$ sample states per initial state,
and each state has $2n$ real numbers.
For the unscented sampling method, the size of the augmented state vector
is $O(dn^{3} + n^{4})$.
There are $n^{2}$ initial states in the operator basis,
$2(2n + d)$ sample states per initial state,
and each state has $2n$ real numbers.
For the $m$\textsuperscript{th}-order derivative method, the size of the augmented state vector
is $O(dmn^{3})$. There are $n^{2}$
initial states in the operator basis, $dm$ state derivatives per initial state,
and each state has $2n$ real numbers.
